Race to the Ballot Box: Will Demographics Decide a Winner This Year?

### 2024-11-01 paper/paper.qmd
Updated paper + added references

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(rstanarm)
library(caret)
library(kableExtra)
library(arrow)

@@

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

@@

# Additional Data Details

>>>>

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: tbl-prediction
#| echo: false
#| 
library(tidyverse)
library(rstanarm)
library(caret)
library(kableExtra)
library(arrow)

@@

## Harris Wins the Swing States and Overcomes the Bradley Effect
Named after Thomas Bradley, a Los Angeles Mayor who underperformed his polls in the 80s, the Bradley Effect is when a candidate does poorly on polls due to his or her gender and/or race [@natesilvergut]. 
From the @tbl-prediction, we can see that Harris wins the majority of the seven swing states,states where "both major parties enjoy similar levels of support among the voting population [@citeSwing]." With these states being the most pivotal toward Harris's path to victory, our prediction of Harris overcoming the Bradley effect, in spite of being a Black woman, bodes well for her in the upcoming election. It is important to realize that due to the nature of the Electoral College, it is possible to not win the popular vote but win the election, and these swing states play a major factor in this regard. Real-world polls, however, show Trump narrowly winning four out of seven swing states (however, the predicted margin of victory still lies within standard error margins of these polls).
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(knitr)

swing_states_data <- data.frame(
  State = c("Pennsylvania", "North Carolina", "Michigan", "Wisconsin", "Georgia", "Arizona", "Nevada"),
  Poll_1 = c(47.7, 47.5, 48.4, 48.5, 47.4, 46.8, 48.0),
  Poll_2 = c(47.7, 47.5, 48.4, 48.5, 47.4, 46.8, 48.0),
  Poll_3 = c(48.3, 48.6, 47.2, 47.8, 49.2, 48.9, 47.9),
  Poll_4 = c(48.3, 48.6, 47.2, 47.8, 49.2, 48.9, 47.9),
  Margin = c("R+ 0.6", "R+ 1.1", "D+ 1.2", "D+ 0.7", "R+ 1.8", "R+ 2.1", "D+ 0.1")
)

kable(swing_states_data, caption = "Swing States Polling Data as of October 31, 2024")
```
It is important to note that the margin of victories in each swing state are within the 2 to 5% margin of error of most polls; it cannot be said for certain who wins these states until Election Day.

## More Urban States Tend to Lean Harris

From our results in @tbl-prediction, we see that Harris has significant leads in states like California and Maryland; from @tbl-urban, notice that these states have a relatively high level of urban population, with 94.2% and 85.6% respectively. One reason for this could be attributed to the higher proportion of educated people in urban areas as opposed to rural ones. This allows people to gain more knowledge in diverse topics such as politics and climate change [@thisnation] and adopt a progressive mindset, which is in line with the ideals of the Democratic Party. In this regard, heavily urban-populated areas tend to be more culturally diverse and open to liberal ideas [@washu_urban_rural]. Moreover, the article states that "people living in rural areas tend to have traditional values and be resistant to new ideas" which goes against the progressive mindset imbibed by Democrats.

## Limitations and Areas for Future Research

@@

# Additional Data Details
```{r}
#| label: tbl-urban
#| echo: false
states_data <- data.frame(
  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", 
            "Colorado", "Connecticut", "Delaware", "Florida", 
            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
            "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
            "Maryland", "Massachusetts", "Michigan", "Minnesota", 
            "Mississippi", "Missouri", "Montana", "Nebraska", 
            "Nevada", "New Hampshire", "New Jersey", "New Mexico",
            "New York", "North Carolina", "North Dakota",
            "Ohio", "Oklahoma", "Oregon", "Pennsylvania",
            "Rhode Island", "South Carolina", "South Dakota",
            "Tennessee", "Texas", "Utah", "Vermont",
            "Virginia", "Washington",  "West Virginia",
            "Wisconsin","Wyoming"),
  urban_percent_2020 = c(57.7, 64.9, 89.3, 55.5, 94.2,
                       86.0, 86.3, 82.6, 91.5,
                       74.1, 86.1, 69.2, 86.9, 71.2,
                       63.2, 72.3, 58.7, 71.5, 38.6,
                       85.6, 91.3, 73.5, 71.9, 46.3,
                       69.5, 53.4, 73.0, 94.1, 58.3,
                       93.8, 74.5, 87.4, 66.7, 61.0,
                       76.3, 64.6, 80.5, 76.5, 91.1, 67.9, 57.2, 66.2, 83.7, 89.8, 35.1,
                       75.6, 83.4, 44.6,
                       67.1,62)
)
kable(states_data, caption = "2020 Percentage of Urban Population for each US State")
```
